---
title: "Week 3"
subtitle: Data Transformations
page-navigation: true
filters: 
  - webr
webr:
  packages:
    - dplyr
    - palmerpenguins
    - janitor
  cell-options:
    editor-font-scale: 0.8
    fig-width: 7
    fig-height: 4.3
    out-width: "100%"
---

## Slides

The slides for this week's lesson are available online as embedded HTML or PDF. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
xaringanExtra::embed_xaringan(
  "https://2026-psyc-259.github.io/3-data-transformations/",
  ratio = "16:9",
  border = "none",
  margin = "0 auto"
)
```

:::{.callout-tip}
Typing <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) in the slide show brings slide navigation controls.
:::

```{r}
#| echo: false
  downloadthis::download_link(
  link = "https://github.com/2026-PSYC-259/3-data-transformations/blob/38c412839cb05dbda2e1c1ed6bbea44b32606b51/slides.pdf?raw=true",
  button_label = "Week 3 Slides PDF",
  button_type = "primary",
  has_icon = TRUE,
  icon = "fa fa-file-alt",
  self_contained = FALSE
  )
```

# Examples

:::{.callout-tip}
After making your changes, Rerun the code by clicking the **{{< fa play title="Play" >}} Run Code** button. If you want to save the code you wrote, you can click *View R History" in the right panel.
:::

## Part 1: Types and classes

We discussed the difference between types and classes in the R language and how to check/convert between them. Try out the brief exercise below during class. 

::: {.panel-tabset}
### Try it out
```{webr-r}
library(dplyr)
ds <- palmerpenguins::penguins
# Which classes are used in the penguins dataset? Which classes?

# Test if the 'species' column is character

# Convert the 'island' column to character

```
### See answer
```{webr-r}
library(dplyr)
ds <- palmerpenguins::penguins
# Which classes are used in the penguins dataset? Which classes?
glimpse(ds) 
#Classes = factor, numeric
class(ds$species)
#Types = integer (including factors), double
typeof(ds$species) 

# Test if the 'species' column is character
is.character(ds$species)

# Convert the 'island' column to character
ds$species <- as.character(ds$species)
glimpse(ds)
```
:::

## Part 2: dplyr

We discussed logical comparisons and using `select`, `filter`, and `slice` to subset data. Try it out in the exercise below.

::: {.panel-tabset}
### Try it out
```{webr-r}
library(dplyr)
ds <- palmerpenguins::penguins

# In one piped statement, filter the dataset to include
# penguins with bills shorter than 39 or longer than 48.
# Assign the results to a new tibble selecting only species and the numeric variables
# Sort the results from largest to smallest bill


# Check if your sorting is correcting by viewing the first 5
# and last 5 rows of the new tibble (in two separate statements)


```
### See answer
```{webr-r}
library(dplyr)
ds <- palmerpenguins::penguins

# In one piped statement, filter the dataset to include
# penguins with bills shorter than 39 or longer than 48.
# Assign the results to a new tibble selecting only species and the numeric variables
# Sort the results from largest to smallest bill
ds_bills <- ds %>% 
  filter(bill_len < 39 | bill_len > 48) %>% 
  select(species, where(is.numeric)) %>% 
  arrange(desc(bill_len))

# Check if your sorting is correcting by viewing the first 5
# and last 5 rows of the new tibble (in two separate statements)
ds_bills %>% slice_head(n = 5)
ds_bills %>% slice_tail(n = 5)
```
:::

We discussed how to `rename`, `group_by`, `mutate`, and `summarize`. Try out the exercise below.

::: {.panel-tabset}
### Try it out
```{webr-r}
library(dplyr)
library(janitor)
ds <- palmerpenguins::penguins

# Change the variable names to lower camel

# Replace NA values for bill variables with the group means provided below
bill_len_mean <- mean(ds$billLen, na.rm = T)
bill_dep_mean <- mean(ds$billDep, na.rm = T)


# Summarize the mean/sd/min/max for billLen and billDep by species/island
# Use across() and starts_with() to avoid typing each calculation

```
### See answer
```{webr-r}
library(dplyr)
library(janitor)
ds <- palmerpenguins::penguins

# Change the variable names to lower camel
ds <- ds %>% clean_names("lower_camel")

# Replace NA values for bill variables with the group means provided below
bill_len_mean <- mean(ds$billLen, na.rm = T)
bill_dep_mean <- mean(ds$billDep, na.rm = T)

ds <- ds %>% 
  mutate(
    billLen = replace_values(billLen, NA ~ bill_len_mean),
    billDep = replace_values(billDep, NA ~ bill_dep_mean),
    )

# Summarize the mean/sd/min/max for billLen and billDep by species/island
# Use across() and starts_with() to avoid typing each calculation
ds %>% 
  group_by(species, island) %>% 
  summarize(
    across(starts_with("bill"), list(mean = mean, sd = sd, min = min, max = max))
  )

```
:::

