---
title: "Week 3"
subtitle: Data Transformations
page-navigation: true
---

## Slides

The slides for this week's lesson are available online as embedded HTML or PDF. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
xaringanExtra::embed_xaringan(
  "https://2026-psyc-259.github.io/3-data-transformations/",
  ratio = "16:9",
  border = "none",
  margin = "0 auto"
)
```

:::{.callout-tip}
Typing <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) in the slide show brings slide navigation controls.
:::

```{r}
#| echo: false
  downloadthis::download_link(
  link = "https://github.com/2026-PSYC-259/3-data-transformations/blob/38c412839cb05dbda2e1c1ed6bbea44b32606b51/slides.pdf?raw=true",
  button_label = "Week 3 Slides PDF",
  button_type = "primary",
  has_icon = TRUE,
  icon = "fa fa-file-alt",
  self_contained = FALSE
  )
```