---
title: "Week 3"
subtitle: Data Transformations
page-navigation: true
---

## Slides

The slides for this week's lesson are available online as embedded HTML or PDF. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
xaringanExtra::embed_xaringan(
  "https://2026-psyc-259.github.io/3-data-transformations/",
  ratio = "16:9",
  border = "none",
  margin = "0 auto"
)
```

:::{.callout-tip}
Typing <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) in the slide show brings slide navigation controls.
:::

```{r}
#| echo: false
  downloadthis::download_link(
  link = "https://github.com/2026-PSYC-259/3-data-transformations/blob/38c412839cb05dbda2e1c1ed6bbea44b32606b51/slides.pdf?raw=true",
  button_label = "Week 3 Slides PDF",
  button_type = "primary",
  has_icon = TRUE,
  icon = "fa fa-file-alt",
  self_contained = FALSE
  )
```

# Examples

## Part 1: Types/classes/logic

::: {.panel-tabset}
### Try it out
```{webr-r}
library(dplyr)
ds <- penguins
# Which classes are used in the penguins dataset? Which classes?

# Test if the 'species' column is character

# Convert the 'island' column to character

```
### See answer
```{webr-r}
library(dplyr)
ds <- penguins
# Which classes are used in the penguins dataset? Which classes?
glimpse(ds) 
#Classes = factor, numeric
class(ds$species)
#Types = integer (including factors), double
typeof(ds$species) 

# Test if the 'species' column is character
is.character(ds$species)

# Convert the 'island' column to character
as.character(ds$species)
```
:::

:::{.callout-tip}
After making your changes, Rerun the code by clicking the **{{< fa play title="Play" >}} Run Code** button. If you want to save the code you wrote, you can click *View R History" in the right panel.
:::
